---
title: "conditionals_loops"
format: html
editor: visual
---

```{r}
library(tidyverse)
```



### Task 1
```{r}
pm2_5 <- 48

if(pm2_5 < 100) {
  print("Low to moderate risk")
} else if(pm2_5 >= 100 & pm2_5 < 150) {
    print("Unhealthy for sensitive groups")
} else if (pm2_5 > 150) {
    print("Health risk present")
}
```
### Task 2
-   Store the string “blue whale” as an object called species. Write an if statement that returns “You found a whale!” if the string “whale” is detected in species, otherwise return nothing. Test by changing the species string & re-running to see output.

```{r}
species <- "blue whale"

if (species == "blue whale") {
  print ("You found a whale!")
} else (print("Not a whale!"))

```

### Task 3
-   Store the base price of a burrito as base_burrito with a value of 6.50. Store main_ingredient with a starting string of “veggie.” Write a statement that will return the price of a burrito based on what a user specifies as “main_ingredient” (either “veggie”, “chicken” or “steak”) given the following:

    A veggie burrito is the cost of a base burrito
    A chicken burrito costs 3.00 more than a base burrito
    A steak burrito costs 3.25 more than a base burrito
    
```{r}
base_burrito <- 6.50
main_ingredient <- "veggie"

#for (value in main_ingredient)

#str_detect(main_ingredient, "veggie", "chicken", "steak")

if (str_detect(main_ingredient,"veggie")) {
  print(base_burrito)
} else if (str_detect(main_ingredient, "chicken")) {
  print(base_burrito + 3)
} else if (str_detect(main_ingredient, "steak")) {
  print(base_burrito + 3.25)

}
# if (main_ingredient == "veggie") {
#   print(base_burrito)
# } else if (main_ingredient == "chicken") {
#   print(base_burrito + 3)
# } else if (main_ingredient == "steak") {
#   print(base_burrito + 3.25)

# switch(main_ingredient,
#        "veggie" = print(base_burrito),
#        "chicken" = print(base_burrito +3),
#        "steak" = print(base_burrito +3.25))

```
 Task 4
 
Create a new vector called fish that contains the values 8, 10, 12, 23 representing counts of different fish types in a fish tank (goldfish, tetras, guppies, and mollies, respectively). Write a for loop that iterates through fish, and returns what proportion of total fish in the tank are that species. Assume that these counts represent all fish in the tank.
 
```{r}
fish <-  c("goldfish" = 8, "tetras" = 10, "guppies" = 12, "mollies" = 23)

for (i in seq_along(fish)){
    proportion_fish <- fish[i] / sum(fish)
      print(proportion_fish)
}
```
Task 5

There is an existing vector in R called month.name that contains all month names (just ry running month.name in the Console to check it out). Write a for loop that iterates over all months in month.name and prints “January is month 1,” “February is month 2”, etc.

Hint: you can index values in the month.name vector just like you would any other vector (e.g., try running month.name[5]).

```{r}
for (i in seq_along(month.name)){
      print(paste(month.name[i], "is month", i))
}
```







